---
title: "GCI Simulator validation"
author: "Daniel Fireman (danielfireman@gmail.com)"
date: "June, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F)
knitr::opts_chunk$set(warning=F)
knitr::opts_chunk$set(cache=F)

require(dplyr)
require(stringr)
require(cowplot)
require(ggplot2)

source("functions.R")

REP <- 1
WARMUP_SECS <- 240
DURATION_SECS <- 0

read.simresult <- function(f) {
  df <- read.csv(f)
  df <- df %>% arrange(timestamp)
  return(list(
    warm=df,
    succ=filter(df, status == 200),
    fail=filter(df, status == 503)
  ))
}

simresult <- function(prefix, n) {
  fname <- paste(prefix, "_1.csv", sep="")
  res <- read.simresult(fname)
  res$succ["expid"] <- 1
  res$succ["id"] <- seq(1,NROW(res$succ))
  if (n == 1) {
    return(res)
  }
  for (i in 2:n) {
    aux <- read.simresult(paste(prefix, "_1.csv", sep=""))
    aux$succ["expid"] <- i
    aux$succ["id"] <- seq(1,NROW(aux$succ))
    res$succ <- rbind(res$succ, aux$succ)
    res$warm <- rbind(res$warn, aux$warn)
    res$fail <- rbind(res$fail, aux$fail)
  }
  return(res)
}
```

```{r}
sim.nogci.1i.nostate <- simresult("sim/sim_output_nogci_sl_300_30_1", REP)
sim.nogci.1i.state <- simresult("sim/sim_output_nogci_sf_300_30_1", REP)
sim.nogci.2i.nostate <- simresult("sim/sim_output_nogci_sl_300_60_2", REP)
sim.nogci.2i.state <- simresult("sim/sim_output_nogci_sf_300_60_2", REP)
sim.nogci.4i.nostate <- simresult("sim/sim_output_nogci_sl_300_120_4", REP)
sim.nogci.4i.state <- simresult("sim/sim_output_nogci_sf_300_120_4", REP)
sim.nogci.cmp <- rbind(
  data.frame("latency"=sim.nogci.1i.nostate$succ$latency, type="OFF", state="Stateless", n="1 Instance"),
  data.frame("latency"=sim.nogci.1i.state$succ$latency, type="OFF", state="Stateful", n="1 Instance"),
  data.frame("latency"=sim.nogci.2i.nostate$succ$latency, type="OFF", state="Stateless", n="2 Instances"),
  data.frame("latency"=sim.nogci.2i.state$succ$latency, type="OFF", state="Stateful", n="2 Instances"),
  data.frame("latency"=sim.nogci.4i.nostate$succ$latency, type="OFF", state="Stateless", n="4 Instances"),
  data.frame("latency"=sim.nogci.4i.state$succ$latency, type="OFF", state="Stateful", n="4 Instances")
)

sim.gci.1i.nostate <- simresult("sim/sim_output_gci_sl_300_30_1", REP)
sim.gci.1i.state <- simresult("sim/sim_output_gci_sf_300_30_1", REP)
sim.gci.2i.nostate <- simresult("sim/sim_output_gci_sl_300_60_2", REP)
sim.gci.2i.state <- simresult("sim/sim_output_gci_sf_300_60_2", REP)
sim.gci.4i.nostate <- simresult("sim/sim_output_gci_sl_300_120_4", REP)
sim.gci.4i.state <- simresult("sim/sim_output_gci_sf_300_120_4", REP)
sim.gci.cmp <- rbind(
  data.frame("latency"=sim.gci.1i.nostate$succ$latency, type="ON", state="Stateless", n="1 Instance"),
  data.frame("latency"=sim.gci.1i.state$succ$latency, type="ON", state="Stateful", n="1 Instance"),
  data.frame("latency"=sim.gci.2i.nostate$succ$latency, type="ON", state="Stateless", n="2 Instances"),
  data.frame("latency"=sim.gci.2i.state$succ$latency, type="ON", state="Stateful", n="2 Instances"),
  data.frame("latency"=sim.gci.4i.nostate$succ$latency, type="ON", state="Stateless", n="4 Instances"),
  data.frame("latency"=sim.gci.4i.state$succ$latency, type="ON", state="Stateful", n="4 Instances")
)

exp.gci.1i.nostate <- accesslog("1i", "gci_T30_M262144_W0", REP, WARMUP_SECS, DURATION_SECS)
exp.gci.1i.state <- accesslog("1i", "gci_T30_M262144_W500", REP, WARMUP_SECS, DURATION_SECS)
exp.gci.2i.nostate <- accesslog("2i", "gci_T60_M262144_W0", REP, WARMUP_SECS, DURATION_SECS)
exp.gci.2i.state <- accesslog("2i", "gci_T60_M262144_W500", REP, WARMUP_SECS, DURATION_SECS)
exp.gci.4i.nostate <- accesslog("4i", "gci_T120_M262144_W0", REP, WARMUP_SECS, DURATION_SECS)
exp.gci.4i.state <- accesslog("4i", "gci_T120_M262144_W500", REP, WARMUP_SECS, DURATION_SECS)
exp.gci.cmp <- rbind(
  data.frame("latency"=exp.gci.1i.nostate$succ$request_time, type="ON", state="Stateless", n="1 Instance"),
  data.frame("latency"=exp.gci.1i.state$succ$request_time, type="ON", state="Stateful", n="1 Instance"),
  data.frame("latency"=exp.gci.2i.nostate$succ$request_time, type="ON", state="Stateless", n="2 Instance"),
  data.frame("latency"=exp.gci.2i.state$succ$request_time, type="ON", state="Stateful", n="2 Instances"),
  data.frame("latency"=exp.gci.4i.nostate$succ$request_time, type="ON", state="Stateless", n="4 Instances"),
  data.frame("latency"=exp.gci.4i.state$succ$request_time, type="ON", state="Stateful", n="4 Instances")
)

exp.nogci.1i.nostate <- accesslog("1i", "nogci_T30_M262144_W0", REP, WARMUP_SECS, DURATION_SECS)
exp.nogci.1i.state <- accesslog("1i", "nogci_T30_M262144_W500", REP, WARMUP_SECS, DURATION_SECS)
exp.nogci.2i.nostate <- accesslog("2i", "nogci_T60_M262144_W0", REP, WARMUP_SECS, DURATION_SECS)
exp.nogci.2i.state <- accesslog("2i", "nogci_T60_M262144_W500", REP, WARMUP_SECS, DURATION_SECS)
exp.nogci.4i.nostate <- accesslog("4i", "nogci_T120_M262144_W0", REP, WARMUP_SECS, DURATION_SECS)
exp.nogci.4i.state <- accesslog("4i", "nogci_T120_M262144_W500", REP, WARMUP_SECS, DURATION_SECS)
exp.nogci.cmp <- rbind(
  data.frame("latency"=exp.nogci.1i.nostate$succ$request_time, type="OFF", state="Stateless", n="1 Instance"),
  data.frame("latency"=exp.nogci.1i.state$succ$request_time, type="OFF", state="Stateful", n="1 Instance"),
  data.frame("latency"=exp.nogci.2i.state$succ$request_time, type="OFF", state="Stateful", n="2 Instances"),
  data.frame("latency"=exp.nogci.2i.state$succ$request_time, type="OFF", state="Stateful", n="2 Instances"),
  data.frame("latency"=exp.nogci.4i.state$succ$request_time, type="OFF", state="Stateful", n="4 Instances"),
  data.frame("latency"=exp.nogci.4i.state$succ$request_time, type="OFF", state="Stateful", n="4 Instances")
)

exp.cmp <- rbind(exp.gci.cmp, exp.nogci.cmp)
sim.cmp <- rbind(sim.gci.cmp, sim.nogci.cmp)
```

It is ideal to have an statistical test to check how close the simulator latency is to the experiment latency. This "closeness" can be measured by a Goodness-of-Fit (GoF) test. We used the two-sample Kolmogorov-Smirnov (KS) test, which checks the null hypothesis: the two samples are drawn from the same distribution. 

The KS test has an issue of being sensitive for large samples. As the data samples used are all large ($10,000+$ datapoints), we applied an approach used in other modelling studies to mitigate this issue i.e. [this](http://ieeexplore.ieee.org/document/5703090/), [this](http://ieeexplore.ieee.org/document/5367061/) and [Marcus' paper](http://ieeexplore.ieee.org/document/6319153/?reload=true) (more at Marcus' paper, including references to goodness-of-fit tests). We selected $1000$ random samples of size $30$ for each fitted data, obtain the p-values for the KS test applied to each sample and then calculate the average p-value.

**1 Instance**
```{r cache=T}
ks.pvalue <- function(e, s) {
  return(mean(bind_rows(
      replicate(
          2000,
          ks.test(sample_n(e, 20)$request_time, sample_n(s, 30)$latency),
          simplify = F),
      .id="Obs")$p.value))
}

ks.1i.table <- as.table(matrix(
  c(
    ks.pvalue(exp.nogci.1i.nostate$succ, sim.nogci.1i.nostate$succ), ks.pvalue(exp.gci.1i.nostate$succ, sim.gci.1i.nostate$succ),
    ks.pvalue(exp.nogci.1i.state$succ, sim.nogci.1i.state$succ), ks.pvalue(exp.gci.1i.state$succ, sim.gci.1i.state$succ)
  ),
  ncol=2, byrow=T)
)
colnames(ks.1i.table) <- c("GCI Off", "GCI On")
rownames(ks.1i.table) <- c("Stateless", "Statefull")
knitr::kable(ks.1i.table)
```

**2 Instances**

```{r cache=T}
ks.2i.table <- as.table(matrix(
  c(
    ks.pvalue(exp.nogci.2i.nostate$succ, sim.nogci.2i.nostate$succ), ks.pvalue(exp.gci.2i.nostate$succ, sim.gci.2i.nostate$succ),
    ks.pvalue(exp.nogci.2i.state$succ, sim.nogci.2i.state$succ), ks.pvalue(exp.gci.2i.state$succ, sim.gci.2i.state$succ)
  ),
  ncol=2, byrow=T)
)
colnames(ks.2i.table) <- c("GCI Off", "GCI On")
rownames(ks.2i.table) <- c("Stateless", "Statefull")
knitr::kable(ks.2i.table)
```

**4 Instances**

```{r cache=T}
ks.4i.table <- as.table(matrix(
  c(
    ks.pvalue(exp.nogci.4i.nostate$succ, sim.nogci.4i.nostate$succ), ks.pvalue(exp.gci.4i.nostate$succ, sim.gci.4i.nostate$succ),
    ks.pvalue(exp.nogci.4i.state$succ, sim.nogci.4i.state$succ), ks.pvalue(exp.gci.4i.state$succ, sim.gci.4i.state$succ)
  ),
  ncol=2, byrow=T)
)
colnames(ks.4i.table) <- c("GCI Off", "GCI On")
rownames(ks.4i.table) <- c("Stateless", "Statefull")
knitr::kable(ks.4i.table)
```

**Conclusion**

As p-values are that high we can not refute the null hyphothesis: one *cannot* claim statistical support for a difference between the simulator and experiment results.


# Appendix

## Kolmogorov-Smirnov Two-Sample Test

More references:

* [KS Test in R](https://stats.stackexchange.com/questions/222294/understanding-kolmogorov-smirnov-test-in-r)

* [KS Test in discrete variables](https://stats.stackexchange.com/questions/48317/kolmogorov-smirnov-with-discrete-data-what-is-proper-use-of-dgofks-test-in-r)

* [KS Test](https://onlinecourses.science.psu.edu/stat414/node/234)

* [KS Test](http://www.physics.csbsju.edu/stats/KS-test.html)