---
title: "GCI Simulator Results"
author: "Daniel Fireman (danielfireman@gmail.com)"
date: "June, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F)
knitr::opts_chunk$set(warning=F)
knitr::opts_chunk$set(cache=F)

require(dplyr)
require(stringr)
require(cowplot)
require(ggplot2)
require(gridExtra)

source("functions.R")

REP <- 2

read.simresult <- function(f) {
  df <- read.csv(f)
  df <- df %>% arrange(timestamp)
  return(list(
    warm=df,
    succ=filter(df, status == 200),
    fail=filter(df, status == 503)
  ))
}

simresult <- function(prefix, n) {
  fname <- paste(prefix, "_1.csv", sep="")
  res <- read.simresult(fname)
  res$succ["expid"] <- 1
  res$succ["id"] <- seq(1,NROW(res$succ))
  if (n == 1) {
    return(res)
  }
  for (i in 2:n) {
    aux <- read.simresult(paste(prefix, "_1.csv", sep=""))
    aux$succ["expid"] <- i
    aux$succ["id"] <- seq(1,NROW(aux$succ))
    res$succ <- rbind(res$succ, aux$succ)
    res$warm <- rbind(res$warn, aux$warn)
    res$fail <- rbind(res$fail, aux$fail)
  }
  return(res)
}
```

```{r}
sim.nogci.1i.nostate <- simresult("sim/sim_output_nogci_sl_300_30_1", REP)
sim.nogci.1i.state <- simresult("sim/sim_output_nogci_sf_300_30_1", REP)
sim.nogci.2i.nostate <- simresult("sim/sim_output_nogci_sl_300_60_2", REP)
sim.nogci.2i.state <- simresult("sim/sim_output_nogci_sf_300_60_2", REP)
sim.nogci.4i.nostate <- simresult("sim/sim_output_nogci_sl_300_120_4", REP)
sim.nogci.4i.state <- simresult("sim/sim_output_nogci_sf_300_120_4", REP)
sim.nogci.8i.nostate <- simresult("sim/sim_output_nogci_sl_300_240_8", REP)
sim.nogci.8i.state <- simresult("sim/sim_output_nogci_sf_300_240_8", REP)
sim.nogci.16i.nostate <- simresult("sim/sim_output_nogci_sl_300_480_16", REP)
sim.nogci.16i.state <- simresult("sim/sim_output_nogci_sf_300_480_16", REP)
sim.nogci.cmp <- rbind(
  data.frame("latency"=sim.nogci.1i.nostate$succ$latency, GCI="OFF", state="Stateless", n="1 Instance"),
  data.frame("latency"=sim.nogci.1i.state$succ$latency, GCI="OFF", state="Stateful", n="1 Instance"),
  data.frame("latency"=sim.nogci.2i.nostate$succ$latency, GCI="OFF", state="Stateless", n="2 Instances"),
  data.frame("latency"=sim.nogci.2i.state$succ$latency, GCI="OFF", state="Stateful", n="2 Instances"),
  data.frame("latency"=sim.nogci.4i.nostate$succ$latency, GCI="OFF", state="Stateless", n="Small (4 Replicas)"),
  data.frame("latency"=sim.nogci.4i.state$succ$latency, GCI="OFF", state="Stateful", n="Small (4 Replicas)"),
  data.frame("latency"=sim.nogci.8i.nostate$succ$latency, GCI="OFF", state="Stateless", n="8 Instances"),
  data.frame("latency"=sim.nogci.8i.state$succ$latency, GCI="OFF", state="Stateful", n="8 Instances"),
  data.frame("latency"=sim.nogci.16i.nostate$succ$latency, GCI="OFF", state="Stateless", n="Large (128 Replicas)"),
  data.frame("latency"=sim.nogci.16i.state$succ$latency, GCI="OFF", state="Stateful", n="Large (128 Replicas)")
)


sim.gci.1i.nostate <- simresult("sim/sim_output_gci_sl_300_30_1", REP)
sim.gci.1i.state <- simresult("sim/sim_output_gci_sf_300_30_1", REP)
sim.gci.2i.nostate <- simresult("sim/sim_output_gci_sl_300_60_2", REP)
sim.gci.2i.state <- simresult("sim/sim_output_gci_sf_300_60_2", REP)
sim.gci.4i.nostate <- simresult("sim/sim_output_gci_sl_300_120_4", REP)
sim.gci.4i.state <- simresult("sim/sim_output_gci_sf_300_120_4", REP)
sim.gci.8i.nostate <- simresult("sim/sim_output_gci_sl_300_240_8", REP)
sim.gci.8i.state <- simresult("sim/sim_output_gci_sf_300_240_8", REP)
sim.gci.16i.nostate <- simresult("sim/sim_output_gci_sl_300_480_16", REP)
sim.gci.16i.state <- simresult("sim/sim_output_gci_sf_300_480_16", REP)
sim.gci.cmp <- rbind(
  data.frame("latency"=sim.gci.1i.nostate$succ$latency, GCI="ON", state="Stateless", n="1 Instance"),
  data.frame("latency"=sim.gci.1i.state$succ$latency, GCI="ON", state="Stateful", n="1 Instance"),
  data.frame("latency"=sim.gci.2i.nostate$succ$latency, GCI="ON", state="Stateless", n="2 Instances"),
  data.frame("latency"=sim.gci.2i.state$succ$latency, GCI="ON", state="Stateful", n="2 Instances"),
  data.frame("latency"=sim.gci.4i.nostate$succ$latency, GCI="ON", state="Stateless", n="Small (4 Replicas)"),
  data.frame("latency"=sim.gci.4i.state$succ$latency, GCI="ON", state="Stateful", n="Small (4 Replicas)"),
  data.frame("latency"=sim.gci.8i.nostate$succ$latency, GCI="ON", state="Stateless", n="8 Instances"),
  data.frame("latency"=sim.gci.8i.state$succ$latency, GCI="ON", state="Stateful", n="8 Instances"),
  data.frame("latency"=sim.gci.16i.nostate$succ$latency, GCI="ON", state="Stateless", n="Large (128 Replicas)"),
  data.frame("latency"=sim.gci.16i.state$succ$latency, GCI="ON", state="Stateful", n="Large (128 Replicas)")
)

cmp <- rbind(sim.gci.cmp, sim.nogci.cmp)
```

```{r}
calc.tp <- function(df) {
  return(NROW(df)/((df[NROW(df),1]-df[1,1])/1000))  
}

calc.tp(sim.nogci.4i.nostate$succ)
calc.tp(sim.gci.4i.nostate$succ)
calc.tp(sim.nogci.4i.state$succ)
calc.tp(sim.gci.4i.state$succ)

calc.tp(sim.nogci.16i.nostate$succ)
calc.tp(sim.gci.16i.nostate$succ)
calc.tp(sim.nogci.16i.state$succ)
calc.tp(sim.gci.16i.state$succ)
```

```{r, cache=T}
rt <- sim.gci.4i.state$succ$latency
m <- median(rt)
p999 <- quantile(rt, c(0.999))

rt.nogci <- sim.nogci.4i.state$succ$latency
m.nogci <- median(rt.nogci)
p999.nogci <- quantile(rt.nogci, c(0.999))

p1 <- ggplot() +
  stat_ecdf(data=sim.gci.4i.state$succ, aes(latency)) +
  ylab("Frequency") +
  xlab("Latency (ms)") +
  theme_bw() +
  geom_vline(xintercept=m, size=0.8, linetype="dotted") +
  geom_text(aes(x=m, label="Median", y=0.4), vjust = 2, angle=90) +
  geom_vline(xintercept=p999, size=0.8, linetype="dotted") +
  geom_text(aes(x=p999, label="99.9th perc.", y=0.4), vjust = 2, angle=90) + 
  geom_segment(data=sim.gci.4i.state$succ, aes(x=m, y=0.1, xend=p999, yend=0.1), arrow = arrow(), size=0.1, colour="darkgray") +
  geom_segment(data=sim.gci.4i.state$succ, aes(x=p999, y=0.1, xend=m, yend=0.1), arrow = arrow(), size=0.1, colour="darkgray") +
  scale_x_continuous(breaks = seq(0, max(rt.nogci), by = 10)) +
  geom_label(aes(x=115, label="GCI On", y=0.5), vjust = 2, angle=0) +
  coord_cartesian(xlim = c(0, max(rt.nogci)))


p2 <- ggplot() +
  stat_ecdf(data=sim.nogci.4i.state$succ, aes(latency)) +
  ylab("Frequency") +
  theme_bw() +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  geom_vline(xintercept=m.nogci, size=0.8, linetype="dotted") +
  geom_text(aes(x=m.nogci, label="Median", y=0.4), vjust = 2, angle=90) +
  geom_vline(xintercept=p999.nogci, size=0.8, linetype="dotted") +
  geom_text(aes(x=p999.nogci, label="99.9th perc.", y=0.4), vjust = 2, angle=90) + 
  geom_segment(data=sim.nogci.4i.state$succ, aes(x=m.nogci, y=0.1, xend=p999.nogci, yend=0.1), arrow = arrow(), size=0.1, colour="darkgray") +
  geom_segment(data=sim.nogci.4i.state$succ, aes(x=p999.nogci, y=0.1, xend=m.nogci, yend=0.1), arrow = arrow(), size=0.1, colour="darkgray") +
  scale_x_continuous(breaks = seq(0, max(rt.nogci), by = 10)) +
  geom_label(aes(x=115, label="GCI Off", y=0.5), vjust = 2, angle=0) +
  coord_cartesian(xlim = c(0, max(rt.nogci)))

ga <- grid.arrange(p2, p1, ncol=1, nrow=2)
ggsave(file="cmp_exp.pdf", ga)
```

```{r}
rt.nogci <- sim.nogci.4i.state$succ$latency
m.nogci <- median(rt.nogci)
p999.nogci <- quantile(rt.nogci, c(0.999))
ggplot() +
  stat_ecdf(data=sim.nogci.4i.state$succ, aes(latency)) +
  xlab("Latency (ms)") +
  ylab("Frequency") +
  theme_bw() +
  geom_vline(xintercept=m.nogci, size=0.8, linetype="dotted") +
  geom_text(aes(x=m.nogci, label="Median", y=0.4), vjust = 2, angle=90) +
  geom_vline(xintercept=p999.nogci, size=0.8, linetype="dotted") +
  geom_text(aes(x=p999.nogci, label="99.9th perc.", y=0.4), vjust = 2, angle=90) + 
  geom_segment(data=sim.nogci.4i.state$succ, aes(x=m.nogci, y=0.1, xend=p999.nogci, yend=0.1), arrow = arrow(), size=0.1, colour="darkgray") +
  geom_segment(data=sim.nogci.4i.state$succ, aes(x=p999.nogci, y=0.1, xend=m.nogci, yend=0.1), arrow = arrow(), size=0.1, colour="darkgray") +
  scale_x_continuous(breaks = seq(0, max(rt.nogci), by = 10)) +
  coord_cartesian(xlim = c(0, max(rt.nogci)))
ggsave("intro.pdf")
```


# Simulation results

### GCI Off

How does a simple service without GCI deal with changes in heap pressure and
keeping state?

```{r}
ggplot(cmp, aes(type, latency)) +
  geom_boxplot() +
  facet_grid(state ~n) +
  scale_y_continuous(breaks=seq(0,max(cmp$latency), 10)) +
  ylab("Latency(ms)") +
  theme(axis.title.x=element_blank())
```
```{r}
ggplot(cmp%>%filter(n == "Small (4 Replicas)" | n == "Large (128 Replicas)"), aes(latency, linetype=GCI)) +
  stat_ecdf() +
  facet_grid(state ~ n) +
  xlab("Latency(ms)") +
  ylab("ECDF at Tail") +
  scale_x_continuous(breaks=seq(0,max(cmp$latency), 30)) +
  coord_cartesian(ylim = c(0.99, 1)) +
  theme_bw() +
  theme(legend.position="top") 
  
ggsave("res_cmp.pdf")
```

```{r}

print.stats.table <- function(df1, df2) {
  ci999.df1 <- ci.fun(df1, p999)
  ci999.diff <- ci.diff.fun(df1, df2, p999)
  ci9999.df1 <- ci.fun(df1, p9999)
  ci9999.diff <- ci.diff.fun(df1, df2, p9999)
  ci99999.df1 <- ci.fun(df1, p99999)
  ci99999.diff <- ci.diff.fun(df1, df2, p99999)
  median.df1 <- ci.fun(df1, median)
  median.diff <- ci.diff.fun(df1, df2, median)
  sd.df1 <- ci.fun(df1, sd)
  sd.diff <- ci.diff.fun(df1, df2, sd)
  print(paste("Median", signif(median.diff$perc[,4:5]/median.df1$perc[,5], digits=4)*100))
  print(paste("SD", signif(sd.diff$perc[,4:5]/sd.df1$perc[,5], digits=4)*100))
  print(paste("99.9", signif(ci999.diff$perc[,4:5]/ci999.df1$perc[,5], digits=4)*100))
  print(paste("99.99", signif(ci9999.diff$perc[,4:5]/ci9999.df1$perc[,5], digits=4)*100))
  print(paste("99.999", signif(ci99999.diff$perc[,4:5]/ci99999.df1$perc[,5], digits=4)*100))
}

SAMPLESIZE <- 1000

print("Small")
print("")
print("Stateless")
print("")
print.stats.table(
  sample(sim.nogci.4i.nostate$succ$latency, SAMPLESIZE),
  sample(sim.gci.4i.nostate$succ$latency, SAMPLESIZE)
)
print("")
print("Stateful")
print("")
print.stats.table(
  sample(sim.nogci.4i.state$succ$latency, SAMPLESIZE),
  sample(sim.gci.4i.state$succ$latency, SAMPLESIZE)
)

print("")
print("Large")
print("")
print("Stateless")
print("")
print.stats.table(
  sample(sim.nogci.16i.nostate$succ$latency, SAMPLESIZE),
  sample(sim.gci.16i.nostate$succ$latency, SAMPLESIZE)
)
print("")
print("Stateful")
print("")
print.stats.table(
  sample(sim.nogci.16i.state$succ$latency, SAMPLESIZE),
  sample(sim.gci.16i.state$succ$latency, SAMPLESIZE)
)
```

```{r}
mean(bind_rows(
      replicate(
          10000,
          ks.test((sim.nogci.4i.nostate$succ %>% sample_n(30))$latency, 
                  (sim.gci.4i.nostate$succ %>% sample_n(30))$latency),
          simplify = F),
      .id="Obs")$p.value)

mean(bind_rows(
      replicate(
          10000,
          ks.test((sim.nogci.16i.nostate$succ %>% sample_n(30))$latency, 
                  (sim.gci.16i.nostate$succ %>% sample_n(30))$latency),
          simplify = F),
      .id="Obs")$p.value)
mean(bind_rows(
      replicate(
          1000,
          ks.test((sim.nogci.4i.state$succ %>% sample_n(30))$latency, 
                  (sim.gci.4i.state$succ %>% sample_n(30))$latency),
          simplify = F))$p.value)

mean(bind_rows(
      replicate(
          10000,
          ks.test((sim.nogci.16i.state$succ %>% sample_n(30))$latency, 
                  (sim.gci.16i.state$succ %>% sample_n(30))$latency),
          simplify = F),
      .id="Obs")$p.value)
```
  
}
```{r}
print("Stateless")
print("")
lat.nogci.16i.nostate <- sample(sim.nogci.16i.nostate$succ$latency, 4000)
lat.gci.16i.nostate <- sample(sim.gci.16i.nostate$succ$latency, 4000)
ci999.nogci.nostate <- ci.quantile(lat.nogci.16i.nostate, 0.999)
ci999.diff.nostate <- ci.diff.quantile(lat.nogci.16i.nostate, lat.gci.16i.nostate, 0.999)
signif(ci999.diff.nostate$basic[,4:5]/ci999.nogci.nostate$basic[,5], digits=4)*100

ci9999.nogci.nostate <- ci.quantile(lat.nogci.16i.nostate, 0.9999)
ci9999.diff.nostate <- ci.diff.quantile(lat.nogci.16i.nostate, lat.gci.16i.nostate, 0.9999)
signif(ci9999.diff.nostate$basic[,4:5]/ci9999.nogci.nostate$basic[,5], digits=4)*100

ci99999.nogci.nostate <- ci.quantile(lat.nogci.16i.nostate, 0.99999)
ci99999.diff.nostate <- ci.diff.quantile(lat.nogci.16i.nostate, lat.gci.16i.nostate, 0.99999)
signif(ci99999.diff.nostate$basic[,4:5]/ci99999.nogci.nostate$basic[,5], digits=4)*100

print("")
print("Stateful")
print("")
lat.nogci.16i.state <- sample(sim.nogci.4i.state$succ$latency, 4000)
lat.gci.16i.state <- sample(sim.gci.4i.state$succ$latency, 4000)
ci999.nogci.state <- ci.quantile(lat.nogci.16i.state, 0.999)
ci999.diff.state <- ci.diff.quantile(lat.nogci.16i.state, lat.gci.16i.state, 0.999)
signif(ci999.diff.state$basic[,4:5]/ci999.nogci.state$basic[,5], digits=4)*100

ci9999.nogci.state <- ci.quantile(lat.nogci.16i.state, 0.9999)
ci9999.diff.state <- ci.diff.quantile(lat.nogci.16i.state, lat.gci.16i.state, 0.9999)
signif(ci9999.diff.state$basic[,4:5]/ci9999.nogci.state$basic[,5], digits=4)*100

ci99999.nogci.state <- ci.quantile(lat.nogci.16i.state, 0.99999)
ci99999.diff.state <- ci.diff.quantile(lat.nogci.16i.state, lat.gci.16i.state, 0.99999)
signif(ci99999.diff.state$basic[,4:5]/ci99999.nogci.state$basic[,5], digits=4)*100
```

#### Latency improvement

How much does the tail latency improve by switching GCI ON?

**1 Instance (%)**
```{r}
t.nogci.1i <- tail.table(
  sim.nogci.1i.nostate$succ$latency,
  sim.nogci.1i.state$succ$latency)
t.gci.1i <- tail.table(
  sim.gci.1i.nostate$succ$latency,
  sim.gci.1i.state$succ$latency)
((t.nogci.1i-t.gci.1i)/t.nogci.1i)*100
```
**2 Instances (%)**
```{r}
t.nogci.2i <- tail.table(
  sim.nogci.2i.nostate$succ$latency,
  sim.nogci.2i.state$succ$latency)
t.gci.2i <- tail.table(
  sim.gci.2i.nostate$succ$latency,
  sim.gci.2i.state$succ$latency)
((t.nogci.2i-t.gci.2i)/t.nogci.2i)*100
```
#### Latency Stddev Improvement (%)

How much does the tail latency variability decrease by switching GCI ON?

```{r}
sd.imp.table <- as.table(matrix(c(
  sd.improvement.perc(sim.nogci.1i.nostate$succ$latency, sim.gci.1i.nostate$succ$latency),
  sd.improvement.perc(sim.nogci.2i.nostate$succ$latency, sim.gci.2i.nostate$succ$latency),
  sd.improvement.perc(sim.nogci.1i.state$succ$latency, sim.gci.1i.state$succ$latency),
  sd.improvement.perc(sim.nogci.2i.state$succ$latency, sim.gci.2i.state$succ$latency)
), ncol=2, byrow=T))
  
colnames(sd.imp.table) <- c("1 Instance", "2 Instances")
rownames(sd.imp.table) <- c("Stateless", "Statefull")
sd.imp.table
```

### Throughput Drop (%)

How much does the throughput drop by switching GCI on?

```{r}
tp.imp.table <- as.table(matrix(c(
  tp.improvement.perc(sim.nogci.1i.nostate$succ, sim.gci.1i.nostate$succ),
  tp.improvement.perc(sim.nogci.2i.nostate$succ, sim.gci.2i.nostate$succ),
  tp.improvement.perc(sim.nogci.1i.state$succ, sim.gci.1i.state$succ),
  tp.improvement.perc(sim.nogci.2i.state$succ, sim.gci.2i.state$succ)
), ncol=2, byrow=T))
  
colnames(tp.imp.table) <- c("1 Instance", "2 Instances")
rownames(tp.imp.table) <- c("Stateless", "Statefull"
)
tp.imp.table
```
