---
title: "GCI Simulator Results"
author: "Daniel Fireman (danielfireman@gmail.com)"
date: "June, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F)
knitr::opts_chunk$set(warning=F)
knitr::opts_chunk$set(cache=F)

require(dplyr)
require(stringr)
require(cowplot)
require(ggplot2)

source("functions.R")

REP <- 2
WARMUP_SECS <- 240
DURATION_SECS <- 0

read.simresult <- function(f) {
  df <- read.csv(f)
  df <- df %>% arrange(timestamp)
  return(list(
    warm=df,
    succ=filter(df, status == 200),
    fail=filter(df, status == 503)
  ))
}

simresult <- function(prefix, n) {
  fname <- paste(prefix, "_1.csv", sep="")
  res <- read.simresult(fname)
  res$succ["expid"] <- 1
  res$succ["id"] <- seq(1,NROW(res$succ))
  if (n == 1) {
    return(res)
  }
  for (i in 2:n) {
    aux <- read.simresult(paste(prefix, "_1.csv", sep=""))
    aux$succ["expid"] <- i
    aux$succ["id"] <- seq(1,NROW(aux$succ))
    res$succ <- rbind(res$succ, aux$succ)
    res$warm <- rbind(res$warn, aux$warn)
    res$fail <- rbind(res$fail, aux$fail)
  }
  return(res)
}
```

```{r}
sim.nogci.1i.nostate <- simresult("sim/sim_output_nogci_sl_300_30_1", REP)
sim.nogci.1i.state <- simresult("sim/sim_output_nogci_sf_300_30_1", REP)
sim.nogci.2i.nostate <- simresult("sim/sim_output_nogci_sl_300_60_2", REP)
sim.nogci.2i.state <- simresult("sim/sim_output_nogci_sf_300_60_2", REP)
sim.nogci.cmp <- rbind(
  data.frame("latency"=sim.nogci.1i.nostate$succ$latency, type="OFF", state="Stateless", n="1 Instance"),
  data.frame("latency"=sim.nogci.1i.state$succ$latency, type="OFF", state="Stateful", n="1 Instance"),
  data.frame("latency"=sim.nogci.2i.nostate$succ$latency, type="OFF", state="Stateless", n="2 Instances"),
  data.frame("latency"=sim.nogci.2i.state$succ$latency, type="OFF", state="Stateful", n="2 Instances")
)

sim.gci.1i.nostate <- simresult("sim/sim_output_gci_sl_300_30_1", REP)
sim.gci.1i.state <- simresult("sim/sim_output_gci_sf_300_30_1", REP)
sim.gci.2i.nostate <- simresult("sim/sim_output_gci_sl_300_60_2", REP)
sim.gci.2i.state <- simresult("sim/sim_output_gci_sf_300_60_2", REP)
sim.gci.cmp <- rbind(
  data.frame("latency"=sim.gci.1i.nostate$succ$latency, type="ON", state="Stateless", n="1 Instance"),
  data.frame("latency"=sim.gci.1i.state$succ$latency, type="ON", state="Stateful", n="1 Instance"),
  data.frame("latency"=sim.gci.2i.nostate$succ$latency, type="ON", state="Stateless", n="2 Instances"),
  data.frame("latency"=sim.gci.2i.state$succ$latency, type="ON", state="Stateful", n="2 Instances")
)

cmp <- rbind(sim.gci.cmp, sim.nogci.cmp)
```

# Simulation results

### GCI Off

How does a simple service without GCI deal with changes in heap pressure and
keeping state?

```{r}
ggplot(cmp, aes(type, latency)) +
  geom_boxplot() +
  facet_grid(state ~n) +
  scale_y_continuous(breaks=seq(0,max(cmp$latency), 10)) +
  ylab("Latency(ms)") +
  theme(axis.title.x=element_blank())
```
```{r}
ggplot(cmp, aes(latency, linetype=type)) +
  stat_ecdf() +
  facet_grid(state ~ n) +
  xlab("Latency(ms)") +
  ylab("ECDF at Tail") +
  scale_x_continuous(breaks=seq(0,max(cmp$latency), 20)) +
  coord_cartesian(ylim = c(0.99, 1)) +
  theme(legend.position="top")
```

#### Latency improvement

How much does the tail latency improve by switching GCI ON?

**1 Instance (%)**
```{r}
t.nogci.1i <- tail.table(
  sim.nogci.1i.nostate$succ$latency,
  sim.nogci.1i.state$succ$latency)
t.gci.1i <- tail.table(
  sim.gci.1i.nostate$succ$latency,
  sim.gci.1i.state$succ$latency)
((t.nogci.1i-t.gci.1i)/t.nogci.1i)*100
```
**2 Instances (%)**
```{r}
t.nogci.2i <- tail.table(
  sim.nogci.2i.nostate$succ$latency,
  sim.nogci.2i.state$succ$latency)
t.gci.2i <- tail.table(
  sim.gci.2i.nostate$succ$latency,
  sim.gci.2i.state$succ$latency)
((t.nogci.2i-t.gci.2i)/t.nogci.2i)*100
```
#### Latency Stddev Improvement (%)

How much does the tail latency variability decrease by switching GCI ON?

```{r}
sd.imp.table <- as.table(matrix(c(
  sd.improvement.perc(sim.nogci.1i.nostate$succ$latency, sim.gci.1i.nostate$succ$latency),
  sd.improvement.perc(sim.nogci.2i.nostate$succ$latency, sim.gci.2i.nostate$succ$latency),
  sd.improvement.perc(sim.nogci.1i.state$succ$latency, sim.gci.1i.state$succ$latency),
  sd.improvement.perc(sim.nogci.2i.state$succ$latency, sim.gci.2i.state$succ$latency)
), ncol=2, byrow=T))
  
colnames(sd.imp.table) <- c("1 Instance", "2 Instances")
rownames(sd.imp.table) <- c("Stateless", "Statefull")
sd.imp.table
```

### Throughput Drop (%)

How much does the throughput drop by switching GCI on?

```{r}
tp.imp.table <- as.table(matrix(c(
  tp.improvement.perc(sim.nogci.1i.nostate$succ, sim.gci.1i.nostate$succ),
  tp.improvement.perc(sim.nogci.2i.nostate$succ, sim.gci.2i.nostate$succ),
  tp.improvement.perc(sim.nogci.1i.state$succ, sim.gci.1i.state$succ),
  tp.improvement.perc(sim.nogci.2i.state$succ, sim.gci.2i.state$succ)
), ncol=2, byrow=T))
  
colnames(tp.imp.table) <- c("1 Instance", "2 Instances")
rownames(tp.imp.table) <- c("Stateless", "Statefull"
)
tp.imp.table
```
